---
header-includes:
- \usepackage{amssymb, amsmath, amsthm}
- \usepackage{tabu}
- \newcommand{\E}{\mathbb{E}}
- \newcommand{\var}{{\rm Var}}
- \newcommand{\N}{\mathcal{N}}
output: pdf_document
---

\noindent \begin{tabu} to \textwidth {@{}X[4 l] @{}X[r]}
  \textbf{HW 4}           & \\ 
  \textbf{MFE 409: Financial Risk Measurement and Management}   & \\ 
  \textbf{Professor Valentin Haddad}         & \\
  \textbf{Group 9}         & \\
  \textbf{Students: Xiahao Wang, Haoxuan Tong, Yuhua Deng, Nupur Solanki}
\end{tabu}


## Qn1

### 1.1

Calculate the Exponential Weighted Moving Average (EWMA)

```{r}
suppressMessages(require(data.table))
suppressMessages(require(lubridate))
suppressMessages(require(zoo))
suppressMessages(require(rugarch))

rm(list=ls())
returns_data <- as.data.table(read.csv("hw3_returns2.csv"))
returns_data[,Date:=mdy(Date)]

# get volatility
returns_data[, volatility := Return^2]
returns_data[, volatility := cumsum(volatility)]
returns_data[, observation := 1]
returns_data[, observation := cumsum(observation)]
returns_data[, volatility := volatility/observation]

lambda <- 0.94

returns_data[, one_minus_lambda := (1 - lambda)* (Return^2)]
returns_data$ewma[1] <- returns_data$one_minus_lambda[1]

size <- length(returns_data$ewma)

for(i in 2:size){
  returns_data$ewma[i] <- lambda * returns_data$ewma[i-1] +  returns_data$one_minus_lambda[i]
}

plot(x=returns_data$Date, y= returns_data$ewma, type="l", xlab = "Date", ylab="Volatility", main = "EWMA")
returns_data[, observation := NULL]
returns_data[, one_minus_lambda := NULL]
```
Next, find the VaR at 99% 
The daily Value at Risk (VaR) is simply a function of the standard deviation or volatility and the desired confidence level. 

Value at Risk (VAR) = sqrt(EWMA predicted Volatility) × z-value of standard normal cumulative distribution

Calculate Daily VaR for both historical and exponential

```{r}
c <- 0.01
z <- qnorm(0.01)
returns_data[, VaR := sqrt(ewma) * z]

data_2005_onwards <- returns_data[Date >= "2015-01-01"]

plot(x = data_2005_onwards$Date, y= data_2005_onwards$Return, type = "l", 
     main="Stock Return against VaRs", xlab = "Date",
     ylab= "Stock Return")
lines(x= data_2005_onwards$Date, y= data_2005_onwards$VaR, col = "blue")
legend("bottomright",legend=c("Return","VaR"),fill=c("black","blue"), cex = 0.8)

Exceptions <- sum(data_2005_onwards[,Return < VaR ])
cat("There are", Exceptions, "exceptions with EWMA")
```




### 1.2 

GARCH(1,1) formula:

$$\sigma^2_t = \gamma V_L + \alpha R^2_{t-1} + \beta \sigma^2_{t-1}$$ 

where $w =  \gamma V_L $

It is EWMA + long-run average

Using Built in Rugarch:

```{r}
library(rugarch)

mymodel <- ugarchspec(variance.model = list(model = "fGARCH", submodel='GARCH', garchOrder = c(1, 1)), mean.model = list(armaOrder = c(0, 0), include.mean = F), 
distribution.model = "norm")

fitVal <- ugarchfit(data =returns_data$Return , spec = mymodel, method="BFGS")
```

Start with the optimal parameters

```{r}
omega <- as.double(fitVal@fit$coef[1])
alpha1 <- as.double(fitVal@fit$coef[2])
beta1 <- as.double(fitVal@fit$coef[3])

cat("omega: ",omega,", alpha: ", alpha1, ", beta1: ", beta1, "\n")

# using the sigma from
garchVol <-fitVal@fit$sigma

returns_data <- cbind(returns_data, garchVol)

returns_data[, GARCHVaR := garchVol * z]

data_2005_onwards <- returns_data[Date >= "2015-01-01"]

plot(x = data_2005_onwards$Date, y= data_2005_onwards$Return, type = "l", 
     main="Stock Return against VaRs Garch (1,1)", xlab = "Date",
     ylab= "Stock Return")
lines(x= data_2005_onwards$Date, y= data_2005_onwards$GARCHVaR, col = "blue")
legend("bottomright",legend=c("Return","VaR Weighted"),fill=c("black","blue"), cex = 0.8)

Except <- sum(data_2005_onwards[,Return < GARCHVaR ])
cat("There are", Except, "exceptions with Garch(1,1)")
```




### 1.3 
Rerun the code from HW3 to get normalized return

```{r}
rm(list=ls())
# from
Q3_returns_data <- as.data.table(read.csv("hw3_returns2.csv"))
Q3_returns_data[,Date:=mdy(Date)]

rolling_win <- 20

# find stddev, mean
Q3_returns_data[, `:=`(volatility = shift(rollapply(Q3_returns_data$Return, rolling_win, sd ,fill=NA, align="right")), 
                       mean = shift(rollapply(Q3_returns_data$Return, 
                       rolling_win, mean ,fill=NA, align="right")))]

# normalized return
Q3_returns_data[,normalized:= (Return - mean)/volatility]

sorted_ret <- sort(Q3_returns_data$normalized)

n <- length(sorted_ret)

sorted_ret_dt <- as.data.table(sorted_ret,col=1)
sorted_ret_dt[, rank := 1]
sorted_ret_dt[, rank := cumsum(rank)]

sorted_ret_dt[, logrank := log(rank) ]
sorted_ret_dt[, logret := log(abs(sorted_ret))]

loss_dt <- sorted_ret_dt[1:(length(sorted_ret_dt$sorted_ret) * 0.05),]

fit <- lm(logrank ~ logret, data=loss_dt)
summary(fit)
plot(x= loss_dt$logret, y= loss_dt$logrank, main="Normalized left tail distribution at 5%", xlab= "log return", ylab="log rank")
abline(a= fit$coefficients[1], b=fit$coefficients[2])
```

Power law:

$$Prob(X > x) = Kx^{-1/\xi}$$

Taking log on both sides 

$$log(Prob(X > x)) = log(K) - \frac{1}{\xi}log(x)$$
In this case, the regression line fits the data pretty well, indicating a that the distribution follows the power law.

Now, estimate the values of $\beta$ and $\xi$ using `optim` function

```{r}
sorted_ret <- sort(na.omit(Q3_returns_data$normalized))

x <-abs(sorted_ret[1:(length(sorted_ret)*0.05)])

u <- last(x)

llike <- function(x, par, u){
  xi <- par[1]
  beta <- par[2]
  n <- length(x)
  ll <- sum(log((1/beta) * (1 + xi * (x - u) / beta) ^ ((-1/xi) - 1)))
  return(-ll)
}

par <- c(0.1,0.1)
gpd.estimate <- optim(par=par, llike, x = x , u = u)

cat("estimates for xi and beta is: ", gpd.estimate$par)
xi <- gpd.estimate$par[1]
beta <- gpd.estimate$par[2]
```

Use the following formula to get VaR:

$$VaR = u + \frac{\beta}{\xi}([\frac{n}{n_u}]^{-\xi} - 1)$$

```{r}
n <- length(sorted_ret)
n_u <- n * 0.05
c <- 0.01

VaR <- u + (beta/xi) *( (n*c/n_u)^(-xi) -1)

cat("For normalized return, the VaR on the last day is: ", VaR,"\n")
```



### 1.4

```{r}
data_2005_onwards <- Q3_returns_data[Date >= "2015-01-01"]

except_normalized <- sum(data_2005_onwards[, normalized < (VaR * -1) ])

cat("Using the VaR calculated from Pareto Distribution, the number of exceptions are: ", except_normalized,"\n")
```

There are 748 data entries from 2015 to 2017. By taking 99% VaR on these observation, we should be seeing (748 * 0.01 = 7.48) 8 exceptions. By using the pareto distribution, we obtained 7 observations. 

In the previous HW3, i obtained 8 exceptions using historical and exponential weighting on normalized return. I would say that pareto distribution is also a good technique to estimate the VaR. 





## Qn2

### 1. What was the broad trading strategy of LTCM ?



### 2. Why did they need so much leverage?

### 3. How did their demise happen? 

### 4. What were the most important issues with their risk management approach?

### 5. How would you manage risk for a fund trying to trade similar strategies?



