---
header-includes:
- \usepackage{amssymb, amsmath, amsthm}
- \usepackage{tabu}
- \newcommand{\E}{\mathbb{E}}
- \newcommand{\var}{{\rm Var}}
- \newcommand{\N}{\mathcal{N}}
output: pdf_document
---

\noindent \begin{tabu} to \textwidth {@{}X[4 l] @{}X[r]}
  \textbf{HW 4}           & \\ 
  \textbf{MFE 409: Financial Risk Measurement and Management}   & \\ 
  \textbf{Professor Valentin Haddad}         & \\
  \textbf{Group 9}         & \\
  \textbf{Students: Xiahao Wang, Haoxuan Tong, Yuhua Deng, Nupur Solanki}
\end{tabu}


## Qn1

### 1.

Calculate the Exponential Weighted Moving Average (EWMA)

```{r}
suppressMessages(require(data.table))
suppressMessages(require(lubridate))
suppressMessages(require(zoo))
suppressMessages(require(rugarch))

rm(list=ls())
returns_data <- as.data.table(read.csv("hw3_returns2.csv"))
returns_data[,Date:=mdy(Date)]

# get volatility
returns_data[, volatility := Return^2]
returns_data[, volatility := cumsum(volatility)]
returns_data[, observation := 1]
returns_data[, observation := cumsum(observation)]
returns_data[, volatility := volatility/observation]

lambda <- 0.94

returns_data[, one_minus_lambda := (1 - lambda)* (Return^2)]
returns_data$ewma[1] <- returns_data$one_minus_lambda[1]

size <- length(returns_data$ewma)

for(i in 2:size){
  returns_data$ewma[i] <- lambda * returns_data$ewma[i-1] +  returns_data$one_minus_lambda[i]
}

plot(x=returns_data$Date, y= returns_data$ewma, type="l", xlab = "Date", ylab="Volatility", main = "EWMA")
returns_data[, observation := NULL]
returns_data[, one_minus_lambda := NULL]
```
Next, find the VaR at 99% 
The daily Value at Risk (VaR) is simply a function of the standard deviation or volatility and the desired confidence level. 

Value at Risk (VAR) = sqrt(EWMA predicted Volatility) × z-value of standard normal cumulative distribution

Calculate Daily VaR for both historical and exponential

```{r}
c <- 0.01
z <- qnorm(0.01)
returns_data[, VaR := sqrt(ewma) * z]

rollingWindow <- 251

cal_hist_VaR <- function(VaRs, c){
  sorted_VaRs <- sort(VaRs)
  Var_point <- ceiling(length(sorted_VaRs) * c)
  var <- sorted_VaRs[Var_point]
  return(var)
}

# calculating exponential VaR
cal_exponential_VaR <- function(VaRs, c){
  n <- length(VaRs)
  weight <- lambda^seq((n-1), 0, -1) * (1-lambda)/(1-lambda^n)
  df <- cbind(weight, VaRs)
  df_sorted <- df[order(VaRs),]
  df_sorted <- cbind(df_sorted, cumsum=cumsum(df_sorted[,1]))
  pos <- which(df_sorted[,3]>c)[1]
  return(df_sorted[pos,2])
}


# calculating historical VaR
returns_data[, historicalVaR := shift(rollapply(returns_data$VaR, rollingWindow, function (VaRs){
  return(cal_hist_VaR(VaRs,c))
}, fill=NA, align="right"))
]

# calculating exponential Weighted VaR
returns_data[, exponentialWeightedVaR := shift(rollapply(returns_data$VaR, rollingWindow, function(VaRs){
  cal_exponential_VaR(VaRs, c)
}, fill=NA, align="right"))]


data_2005_onwards <- returns_data[Date >= "2015-01-01"]

plot(x = data_2005_onwards$Date, y= data_2005_onwards$Return, type = "l", 
     main="Stock Return against VaRs", xlab = "Date",
     ylab= "Stock Return")
lines(x= data_2005_onwards$Date, y= data_2005_onwards$historicalVaR, col = "blue")
lines(x= data_2005_onwards$Date, y= data_2005_onwards$exponentialWeightedVaR, col = "green")
legend("bottomright",legend=c("Historical","Exponential Weighted"),fill=c("blue","green"), cex = 0.8)

historicalExcept <- sum(data_2005_onwards[,Return < historicalVaR ])
exExcept <-sum(data_2005_onwards[,Return < exponentialWeightedVaR ])

cat("There are", historicalExcept, "exceptions using the historical VaR while", exExcept, "using the exponential Weighted VaR")
```

### 2. 

GARCH(1,1) formula:

$$\sigma^2_t = \gamma V_L + \alpha R^2_{t-1} + \beta \sigma^2_{t-1}$$ 

where $w =  \gamma V_L $

It is EWMA + long-run average

Using Built in Rugarch:

```{r}
library(rugarch)

mymodel <- ugarchspec(variance.model = list(model = "fGARCH", submodel='GARCH', garchOrder = c(1, 1)), mean.model = list(armaOrder = c(0, 0), include.mean = F), 
distribution.model = "norm")

fitVal <- ugarchfit(data =returns_data$Return , spec = mymodel, method="BFGS")
```

Start with the optimal parameters

```{r}
omega <- as.double(fitVal@fit$coef[1])
alpha1 <- as.double(fitVal@fit$coef[2])
beta1 <- as.double(fitVal@fit$coef[3])

# using the sigma from
garchVol <-fitVal@fit$sigma

returns_data <- cbind(returns_data, garchVol)

returns_data[, GARCHVaR := garchVol * z]

c <- 0.01
rollingWindow <- 251
# calculating historical VaR
returns_data[, hVaRGarch := shift(rollapply(returns_data$GARCHVaR, rollingWindow, function (VaRs){
  return(cal_hist_VaR(VaRs,c))
}, fill=NA, align="right"))
]

# calculating exponential Weighted VaR
returns_data[, ewVaRGarch := shift(rollapply(returns_data$GARCHVaR, rollingWindow, function(VaRs){
  cal_exponential_VaR(VaRs, c)
}, fill=NA, align="right"))]


data_2005_onwards <- returns_data[Date >= "2015-01-01"]

plot(x = data_2005_onwards$Date, y= data_2005_onwards$Return, type = "l", 
     main="Stock Return against VaRs Garch (1,1)", xlab = "Date",
     ylab= "Stock Return")
lines(x= data_2005_onwards$Date, y= data_2005_onwards$hVaRGarch, col = "blue")
lines(x= data_2005_onwards$Date, y= data_2005_onwards$ewVaRGarch, col = "green")
legend("bottomright",legend=c("Historical","Exponential Weighted"),fill=c("blue","green"), cex = 0.8)

historicalExcept <- sum(data_2005_onwards[,Return < hVaRGarch ])
exExcept <-sum(data_2005_onwards[,Return < ewVaRGarch ])

cat("There are", historicalExcept, "exceptions using the historical VaR while", exExcept, "using the exponential Weighted VaR")

```

